Figure two

### 2025-12-09 draftFigure2.Rmd
reorganizing
adding weighted category paper count also remove duplicate rows in dataset for FRA and CYP

Redo figure 2 with the 'share of papers mentioning a topic' instead, so the percentages would go above 100% which is ok. Also for the other panels, calculate the percentages per subgroups that mention,
which will also make each 'column' going above 100%

@@

@@

## surveys: drop items for which Inclusion == 0
surveys <- surveys[which(surveys$Inclusion == 1),]

@@

surveys_long <- surveys %>%
  mutate(item_id = paste0(iso3, id, year)) %>% 
  distinct(item_id, category) %>%     # collapse multiple rows for same ID–category pair
  count(category, name = "n") %>% 
  mutate(share = n / length(unique(paste0(surveys$iso3, surveys$id)))) %>%
  rename(Topic = category)

@@

@@

>>>>

@@

papers <- papers %>% 
  filter(!is.na(iso3))

surveys <- surveys %>% 
  filter(!is.na(iso3))

dataSummary <- dataSummary %>% 
  filter(!is.na(iso3))

@@

## surveys: drop items for which Inclusion == 0
surveys <- surveys[which(surveys$Inclusion == 1),] # done in the data prep section

@@

surveys_long <- surveys %>%
  mutate(item_id = paste0(iso3, id, year)) %>% 
  distinct(item_id, category) %>%     # collapse multiple rows for same ID–category pair
  count(category, name = "n") %>% 
  mutate(share = n / length(unique(paste0(surveys$iso3, surveys$id, surveys$year)))) %>%
  rename(Topic = category)

@@

# Visualizations by income group, with proportional weighting

```{r visualisation income weighting}

## 1. Compute country-weights per paper x income group -----------------------

# For each paper, how many distinct countries are there in total,
# and how many of those are in each income group?
paper_country_weights <- papers %>%
  distinct(ID, country, Income.group) %>%  # ignore category here
  group_by(ID) %>%
  mutate(n_countries_total = n_distinct(country)) %>%
  group_by(ID, Income.group, n_countries_total) %>%
  summarise(n_countries_group = n_distinct(country), .groups = "drop") %>%
  mutate(country_weight = n_countries_group / n_countries_total)
# Example: if a paper has 2 countries (1 low, 1 high) then
# country_weight = 0.5 for low income and 0.5 for high income.

## 2. Attach weights back to the (paper x income group x category) rows ------

papers_weighted <- papers %>%
  left_join(
    paper_country_weights %>%
      dplyr::select(ID, Income.group, country_weight),
    by = c("ID", "Income.group")
  )

## 3. Compute weighted category shares within each income group --------------

topics_data <- papers_weighted %>%
  filter(!is.na(Income.group), !is.na(category)) %>%
  group_by(Income.group, category) %>%
  summarise(
    weighted_n = sum(country_weight, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  rename(Topic = category) %>%
  mutate(
    Income.group = factor(
      Income.group,
      levels = c("Low income", "Lower middle income", "Upper middle income", "High income")
    ),
    Topic = factor(
      Topic,
      levels = rev(c("Perceptions", "Behavior", "Priority", "Policy", "Health"))
    )
  ) %>%
  group_by(Income.group) %>%
  mutate(
    share = weighted_n / sum(weighted_n),
    total = sum(weighted_n)        # weighted total "papers" per income group
  ) %>%
  ungroup()

## 4. Data for the top "total papers" row -----------------------------------

totals_data <- topics_data %>%
  group_by(Income.group) %>%
  summarise(total = round(first(total)), .groups = "drop") %>%
  mutate(Topic = "Total papers")   # so it can align in the same plot if needed

## 5. Plot (heatmap) --------------------------------------------------------

pdf("plots/Figure2_category_paper_by_income_weight.pdf", width = 12, height = 10)
ggplot() +
  # --- top totals row ---
  geom_tile(data = totals_data,
            aes(x = Income.group, y = "Total papers", fill = total),
            color = "white", linewidth = 0.8) +
  geom_text(
    data = totals_data,
    aes(x = Income.group, y = "Total papers",
        label = paste0("n=", total),
        color = "black"),
    size = 7
  ) +
  scale_fill_gradient(
    name = "Total papers",
    low = "#F6DBA7", high = "#F7B32B",
    breaks = range(totals_data$total),
    labels = range(totals_data$total)
  ) +
  scale_color_identity() +

@@

# Visualizations by exposure group, with proportional weighting

```{r visualisation categoryPollution weighting}

## 1. Compute country-weights per paper x categoryPollution -----------------------

# For each paper, how many distinct countries are there in total,
# and how many of those are in each income group?
paper_country_weights <- papers %>%
  distinct(ID, country, categoryPollution) %>%  # ignore category here
  group_by(ID) %>%
  mutate(n_countries_total = n_distinct(country)) %>%
  group_by(ID, categoryPollution, n_countries_total) %>%
  summarise(n_countries_group = n_distinct(country), .groups = "drop") %>%
  mutate(country_weight = n_countries_group / n_countries_total)
# Example: if a paper has 2 countries (1 low, 1 high) then
# country_weight = 0.5 for low income and 0.5 for high income.

## 2. Attach weights back to the (paper x categoryPollution x category) rows ------

papers_weighted <- papers %>%
  left_join(
    paper_country_weights %>%
      dplyr::select(ID, categoryPollution, country_weight),
    by = c("ID", "categoryPollution")
  )

## 3. Compute weighted category shares within each categoryPollution --------------

topics_data <- papers_weighted %>%
  filter(!is.na(categoryPollution), !is.na(category)) %>%
  group_by(categoryPollution, category) %>%
  summarise(
    weighted_n = sum(country_weight, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  rename(Topic = category) %>%
  mutate(
    Topic = factor(
      Topic,
      levels = rev(c("Perceptions", "Behavior", "Priority", "Policy", "Health"))
    )
  ) %>%
  group_by(categoryPollution) %>%
  mutate(
    share = weighted_n / sum(weighted_n),
    total = sum(weighted_n)        # weighted total "papers" per categoryPollution
  ) %>%
  ungroup()

## 4. Data for the top "total papers" row -----------------------------------

totals_data <- topics_data %>%
  group_by(categoryPollution) %>%
  summarise(total = round(first(total)), .groups = "drop") %>%
  mutate(Topic = "Total papers")   # so it can align in the same plot if needed

## 5. Plot (heatmap) --------------------------------------------------------

pdf("plots/Figure2_category_paper_by_exposure_weight.pdf", width = 12, height = 10)
ggplot() +
  # --- top totals row ---
  geom_tile(data = totals_data,
            aes(x = categoryPollution, y = "Total papers", fill = total),
            color = "white", linewidth = 0.8) +
  geom_text(
    data = totals_data,
    aes(x = categoryPollution, y = "Total papers",
        label = paste0("n=", total),
        color = "black"),
    size = 7
  ) +
  scale_fill_gradient(
    name = "Total papers",
    low = "#F6DBA7", high = "#F7B32B",
    breaks = range(totals_data$total),
    labels = range(totals_data$total)
  ) +
  scale_color_identity() +

  # --- main topic rows ---
  geom_tile(data = topics_data,
            aes(x = categoryPollution, y = Topic, fill = share),
            color = "white", linewidth = 0.8) +
  geom_text(
    data = topics_data,
    aes(x = categoryPollution, y = Topic,
        label = scales::percent(share, accuracy = 1),
        # label = paste0(
        #     scales::percent(share, accuracy = 1), 
        #     "\n(", n_papers, ")"
        #   ),
        color = ifelse(share > 0.25, "white", "black")),
    size = 7
  ) +
  scale_fill_gradient(
    name = "Share of papers\nwithin exposure group ",
    low = "#f7f4f9", high = "#8B475D",
    labels = scales::percent,
    breaks = c(min(topics_data$share), max(topics_data$share))
  ) +
  scale_color_identity() +
  labs(x = "Exposure group", y = NULL) +
  theme_minimal(base_size = 19) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1, size = 22),
    axis.text.y = element_text(size = 22),
    axis.title = element_text(size = 22),
    legend.position = "top"
  )
dev.off()
